### 2020-06-10 dissertation/final/Chapters/appendix1.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% appendix1.tex
%% NOVA thesis document file
%%
%% Chapter with example of appendix with a short dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Apendíce 1: Gramática}
\label{grammar}
\begin{lstlisting}
main <- body EOF;

body <- statement ( ";" statement )* _ ";"? _
    / ""
    ;

// Statements
statement <- _ ( var_declaration / voice_declaration / function_declaration / for_loop_statement / while_loop_statement / if_statement / expression ) _;

var_declaration <- "$" namespaced _ "=" _ expression;

voice_declaration <- ":" identifier _ "=" _ voice_declaration_body;

voice_declaration_body <- integer
                    / "(" _ function_parameters _ ")"
                    / ":" identifier _ "(" _ function_parameters _ ")"
                    ;

function_declaration <- "fun" _ namespaced _ "(" _ arguments? _ ")" _ "{" body "}";

arguments <- single_argument ( _ ";" _ single_argument )*;

single_argument <- single_argument_expr / single_argument_ref / single_argument_eval;

single_argument_expr <- "expr" _ "$" identifier;

single_argument_ref <- "ref" _ "$" identifier;

single_argument_eval <- "$" identifier;

for_loop_statement <- "for" _ "$" namespaced _ "in" _ value_expression _ ".." _ value_expression _ "{" _ body? _ "}";

while_loop_statement <- "while" _ expression _ "{" _ body _ "}";

if_statement <- "if" _ expression _ "{" _ body _ "}" _ "else" _ "{" _ body _ "}"
            / "if" _ expression _ "{" _ body _ "}";

// BEGIN Keyboard
keyboard_declaration <- "@keyboard" (_ alphanumeric)* ( _ group )? _ "{" ( _ keyboard_shortcut _  ";" )* _ "}";

keyboard_shortcut <- alphanumeric (_"+"_ alphanumeric)* (_ alphanumeric)* _ ":" _ expression
                / string_value (_ alphanumeric)* _ ":" _ expression
                / "[" _ list_comprehension _ "]" (alphanumeric _)* _ ":" _ expression
                / "[" _ value_expression _ "]" (alphanumeric _)* _ ":" _ expression
                ;

list_comprehension <- expression "for" _ "$" _ namespaced _ "in" _ value_expression _ ".." _ value_expression
                / expression "for" _ "$" _ namespaced _ "in" _ value_expression _ ".." _ value_expression _ "if" value_expression
                ;
// END Keyboard

expression <- e music_expression;

music_expression <- sequence ( _ "|" _ sequence )*;

sequence <- value_expression ( _ value_expression)*;

group <- "(" _ expression _ ")";

block <- "{" _ body _ "}";

note <- note_accidental _ note_pitch ( _ chord_suffix )? ( _ note_value )?;

chord_suffix <- 'M' / 'm';

variable <- "$" namespaced;

function <- namespaced "(" _ function_parameters? _ ")";

function_parameters <- expression ( _ ";" _ expression )*;

chord <- "[" _ note ( _ note )* _ "]";

rest <- "r" ( _ note_value )?;

note_value
<- "/" _ integer
    / integer _ "/" _ integer
    / integer
    ;

note_accidental <- "^" / "^^" / "__" / "_" / "";

note_pitch
<- r"[cdefgab]" "'"*
    / r"[CDEFGAB]" ","*
    ;

modifier
<- r"[tT]" _ integer
    / r"[vV]" _ integer
    / r"[lL]" _ note_value
    / r"[sS]" _ integer _ "/" _ integer
    / r"[sS]" _ integer
    / r"[oO]" _ integer
    ;

instrument_modifier <- ":" ( identifier / "?" ) _ sequence;

value_expression <- expression expr_binary_op expression
                / expr_unary_op expression
                / string_value 
                / number_value 
                / bool_value 
                / none_value 
                / variable 
                / function 
                / keyboard_declaration 
                / group 
                / block 
                / chord 
                / note 
                / rest 
                / modifier 
                / instrument_modifier
                ;

exp_binary_op <- "and" / "or" / ">=" / ">" / "==" / "!=" / "<=" / "<" / r"[+\-*/]";

expr_unary_op <- "not";

string_value <- double_string / single_string;

double_string <- "\"" double_string_char* "\"";

double_string_char 
<- "\\\""
    / "\\\\"
    / r"[^\"]"
    ;

single_string <- "'" single_string_char* "'";

single_string_char
<- "\\'"
    / "\\\\"
    / r"[^']"
    ;

number_value <- float / integer;

bool_value <- "true" / "false";

none_value <- "none";

float <- r"[0-9]+\.[0-9]+";

integer <- r"[0-9]+";

namespaced <- ( identifier "\\" )* identifier;

identifier <- r"[a-zA-Z][a-zA-Z0-9\_]*";

alphanumeric <- r"[a-zA-Z0-9\_]*";

_ <- r"[ \t\r\n]*";

e <- "";

comment <- _ r"#[^\n]*";
\end{lstlisting}