### 2020-07-04 dissertation/final/Chapters/glossary.tex
Update dissertation

%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% glossary.tex
%% NOVA thesis document file
%%
%% Glossary definition
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newglossaryentry{computer} {
	name={computer}, 
	description={An electronic device which is capable of receiving information (data) in a particular form and of performing a sequence of operations in accordance with a predetermined but variable set of procedural instructions (program) to produce a result in the form of information or signals.}
}

\newglossaryentry{ligaduras} {
	name={ligaduras}, 
	description={Símbolo colocado nas partituras músicais para ligar notas seguidas com o mesmo \textit{pitch} em diferentes compassos. As durações das notas são somadas e elas são tocadas como uma só nota.}
}

\loadglsentries{example-glossaries-long.tex}
\glsaddall