### 2020-08-11 dissertation/final/Chapters/appendix1.tex
Add Lark section and Future Work

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% appendix1.tex
%% NOVA thesis document file
%%
%% Chapter with example of appendix with a short dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Apendíce 1: Gramática PEG}
\label{grammar}

\lstinputlisting[caption={Gramática PEG da linguagem Musikla}]{../../code/musikla/musikla/parser/grammar.peg}